Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-02-12 Multiresolution_framework/ltx/manuscript.tex
edits to appendix

As mentioned earlier, the E-step for state-space models is equivalent to the smoothing problem. We use the RTS smoother, given in Table~\ref{alg:MRA-RTS} for completeness, in order to infer the states of the system. The RTS smoother uses the Kalman filter \citep{Kalman1960} in forward pass (Steps~1 and 2 in Table~\ref{alg:MRA-RTS}), along with a backward iteration to calculate its outputs (Steps~3 and 4 in Table~\ref{alg:MRA-RTS}): smoothed states $\hat{\mathbf x}^b_t$ and covariances, $\mathbf P^b_t=\mathrm{cov}(\mathbf{x}_t)$. Following \cite{Gibsona2005}, an additional backwards recursion is also included (Step~5 in Table~\ref{alg:MRA-RTS})	to calculate cross-covariance matrix, $\mathbf M_t=\mathrm{cov}(\mathbf{x}_{t},\mathbf{x}_{t+1})$, which is also required in forming the lower bound to be maximized in the M-step. Note that the following quantities from the forward iteration should be stored when implementing the RTS smoother: predicted and corrected estimates for both state estimates ($\hat{\mathbf{x}}_t^{f-}$, $\hat{\mathbf{x}}_t^{f}$) and covariance matrices ($\mathbf P_t^{f-}$,$\mathbf P_t^f$). Also, the final value of the Kalman gain, $\mathcal K_T$, should be stored in order to initialize the backward cross-covariance computation. 

@@

For the linear state-space model given by Eqs.~\eqref{eq:StateEquation} and \eqref{eq:ReducedObservationEquation}, the bound defined by Eq.~\eqref{eq:Bishopbound} is a quadratic of the form (see \ref{ap:QDerivation} for derivation) 
\begin{equation}\label{eq:MRA-QCompact}
\mathcal Q\left(\boldsymbol \theta,\boldsymbol\theta'\right)=\beta+2T_s\varsigma\left(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1\right)\boldsymbol\theta-T_s\varsigma\boldsymbol\theta^\top\boldsymbol\Upsilon\boldsymbol\theta,
\end{equation}
where $\beta$ is constant with respect to $\boldsymbol\theta$. Equation~\eqref{eq:MRA-QCompact} is maximum at
\begin{align}\label{eq:MRA-thetahat}
\boldsymbol \theta= \boldsymbol\Upsilon^{-\top}(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top,
\end{align}
where
\begin{align}
\boldsymbol\upsilon_0&=\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_0]_{i,j}[\boldsymbol\Gamma_1]_{j,i}\label{eq:epsilon0} \\
\boldsymbol\upsilon_1&=\xi\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\boldsymbol\Gamma_1]^{j,i} \label{eq:epsilon1}\\
\boldsymbol\Upsilon&=T_s\varsigma\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\boldsymbol\Gamma_2]^{j,i},\label{eq:Epsilon}
\end{align} 
where $[\cdot]_{p,q}$ denotes the $\left(p,q\right)$-element of the matrix, and $ [\cdot]^{p,q}$ denotes the $\left(p,q\right)$-block of the block matrix and where 
\begin{align}
\left[ \Gamma_1\right]^{j,i} &=\sum_{k=1}^{n_x}\left[ \boldsymbol\Sigma_w^{-1}\boldsymbol\Lambda_x^{-1}\right]_{jk} \left[ \mathbf U\right]^{k,i},\\
\left[ \Gamma_2\right] ^{j,i}&=\sum_{k,m=1}^{n_x}[\mathbf U^{\top}]^{jk} \left[\boldsymbol\Lambda_x^{-1}\boldsymbol\Sigma_w^{-1}\boldsymbol\Lambda_x^{-1} \right]_{km}[\mathbf U]^{mi}.
\end{align}
Each $ 1 \times n_{\theta}$ block of the $n_x \times n_x n_{\theta}$ block matrix $\mathbf U$ is 
\begin{align}\label{eq:Uij}
\left[ \mathbf U\right] ^{k,i}&=\int_{\boldsymbol \Omega}\left[\boldsymbol\mu(r) \right]_k \left[\int_{\boldsymbol\Omega} \boldsymbol\mu\left(r'\right)\boldsymbol \lambda^\top \left(r-r'\right) \mathrm{d}r'\right]_{i:} \mathrm{d}r,
\end{align}
where $[\cdot]_{i:} $ denotes the $i^{th}$ row.  Each element of $\boldsymbol\Lambda_{\theta}$ can be calculated using
\begin{equation}\label{eq:MRA-LambdaThetaElements}
\left[ \boldsymbol\Lambda_{\theta}\right] _{k,i}=\left[ \mathbf U\right]^{k,i}\boldsymbol\theta.
\end{equation}
The matrices $\boldsymbol\Xi_0$ and $\boldsymbol\Xi_1$ in Eqs.~\eqref{eq:epsilon0}-\eqref{eq:Epsilon} are calculated using the RTS smoother outputs: $\hat{\mathbf x}_t$, covariance, $\mathbf P_t^b$, and cross-covariance matrix, $\mathbf M_t$ \citep{Gibsona2005}, 
\begin{align}
\boldsymbol\Xi_0&=\sum_{t=0}^{T-1}\left(\mathbf M_{t+1}+\mathbf{\hat x}_t\mathbf{\hat x}_{t+1}^\top\right) \label{eq:Xi0} \\
 \boldsymbol\Xi_1&=\sum_{t=0}^{T-1}\left(\mathbf P_t+\mathbf{\hat x}_t\mathbf{\hat x}_t^\top\right).  \label{eq:Xi1}
\end{align} 
The second derivative of the $\mathcal Q$-function is given by
\begin{equation}
\frac{\partial^2\mathcal Q}{\partial\boldsymbol\theta^2}=-2T_s\varsigma\boldsymbol\Upsilon.
\end{equation}
The matrix $\boldsymbol\Upsilon$ is positive definite if $n_x^2\times n_{\theta}$ matrix $\mathrm{vec}(\mathbf U)$ is of rank $n_{\theta}$ (\citep{Dewar2009}, Lemma 3), where $\mathrm{vec}(\cdot)$ denotes the vectorization operator. Under this condition $\boldsymbol\Upsilon$ is  invertible and the second derivative is negative definite representing a maximum of the $\mathcal Q$-function.

\subsection{Stopping rule}
The stopping criterion is usually associated with either the change in the parameter estimates or the log-likelihood variation \citep{McLachlan1997}. The stopping rule adopted herein is
\begin{equation}
 \left(\parallel \mathbf{A} \parallel_{F}^{(i)}-\parallel \mathbf{A} \parallel_{F}^{(i-1)}\right)<\epsilon,
 \end{equation}
 where $\epsilon$ is a threshold value and $\parallel \mathbf{A} \parallel_{F}^{(i)}$ and $ \parallel \mathbf{A} \parallel_{F}^{(i-1)}$ are the Frobenius norms  of the successive estimates of $\mathbf{A} $ matrices which is defined by \citet{Meyer2000}
 \begin{equation}
  \parallel \mathbf{A} \parallel_{F}=\sqrt{\sum_{i,j=1}^{n_x}\mid a_{i,j} \mid^2}=\sqrt{\mathrm{tr} (\mathbf A^{\top}\mathbf A)}.
 \end{equation}
\subsection{Computational complexity} 
All $n_x^2$ blocks of $\mathbf U$, $\boldsymbol\Gamma_1$ and $\boldsymbol\Gamma_2$  can be computed as a one-off before the commencement of the EM iterations, which increases the speed of the M-step significantly compared to the implementation of \citet{Dewar2009}. The complexity of each step of the algorithm is summarized in Table~\ref{table:MRA-ComputationalComplexity} where for ease of comparison the computational complexity of the method of \citet{Dewar2009} is also identified. It can be seen from the table that the state dimension is an important factor at each iteration of the estimation algorithm, emphasizing the advantage of the proposed algorithm compared to the method proposed in \citet{Dewar2009}. This also suggests the use of an efficient algorithm to choose the minimum number of basis functions (states of the system) while achieving the desired details of the underlying field.  
\begin {table}
\begin{center}
\scalebox{1}{\begin{tabular}{lllc}
\hline \hline
Variable &Equation&Order&Order obtained from \citet{Dewar2009}\\ \hline\\
$\Xi_0, \Xi_1$&Eqs.~\eqref{eq:Xi0} and \eqref{eq:Xi1}&$O(Tn_x^2)$& \----\\
$\boldsymbol\upsilon_0, \boldsymbol\upsilon_1$&Eqs.~\eqref{eq:epsilon0} and \eqref{eq:epsilon1} &$O(n_x^2n_{\theta})$ &$O(n_x^2n_{\theta})+O(n_x^3)$ \\
$\boldsymbol\Upsilon$&Eq.~\eqref{eq:Epsilon}&$O(n_x^2n_{\theta}^2)$&$O(n_x^4n_{\theta}^2)$\\
$\mathbf A(\boldsymbol\theta)$&Eq.~\eqref{eq:A_theta}&$O(n_x^3)$&\----\\
$\boldsymbol\Lambda_{\theta}$&Eq.~\eqref{eq:MRA-LambdaThetaElements}&$O(n_x^2n_{\theta})$&\----\\
$\hat{\boldsymbol\theta} $&Eq.~\eqref{eq:MRA-thetahat}&$O(n_{\theta}^3)$&\----\\
\hline \hline
\end{tabular}}
 \caption {{\bf The Computational complexity of the estimation algorithm}. A comparison between the algorithm proposed herein and that of \citet{Dewar2009} is also provided. "\----'' indicate equal complexities for both algorithms.} 
\label{table:MRA-ComputationalComplexity}
\end{center}
\end {table}

@@

where $[\cdot]_{p,q}$ denotes the $\left(p,q\right)$-element of the matrix. Each element of $\boldsymbol\Lambda_{\theta}$ can be calculated using
\begin{equation}\label{eq:MRA-LambdaThetaElements}

@@

where $ [\cdot]^{p,q}$ denotes the $\left(p,q\right)$-block of the block matrix. Each $ 1 \times n_{\theta}$ block of the $n_x \times n_x n_{\theta}$ block matrix $\mathbf U$ is 
\begin{align}\label{eq:Uij}

@@

where $[\cdot]_{i:} $ denotes the $i^{th}$ row.  By substituting for $\left[ \boldsymbol\Lambda_{\theta}\right] _{k,i}$ from Eq.~\eqref{eq:MRA-LambdaThetaElements} the trace given in Eq.~\eqref{eq:MRA-trace1} can be written as

@@

\begin{equation}\label{eq:appepsilon0}
\boldsymbol\upsilon_0=\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_0]_{i,j}[\boldsymbol\Gamma_1]_{j,i},
\begin{equation}\label{eq:appepsilon1}
\boldsymbol\upsilon_1=\xi\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\boldsymbol\Gamma_1]^{j,i},
and
\begin{equation}\label{eq:appEpsilon}
\boldsymbol\Upsilon=T_s\varsigma\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\boldsymbol\Gamma_2]^{j,i}.
By differentiating the $\mathcal Q$-function with respect to $\boldsymbol\theta$ we have
\begin{align}\label{eq:MRA-QDerivative}
\frac{\partial \mathcal Q}{\partial \boldsymbol\theta}&=2T_s\varsigma(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top-T_s\varsigma(\boldsymbol\Upsilon^\top+\boldsymbol\Upsilon)\boldsymbol\theta \nonumber \\
&=2T_s\varsigma(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top-2T_s\varsigma\boldsymbol\Upsilon^\top\boldsymbol\theta.
Equating Eq.~\eqref{eq:MRA-QDerivative} to $\mathbf 0$ vector yields
\begin{align}\label{eq:MRA-thetahat}
\boldsymbol \theta= \boldsymbol\Upsilon^{-\top}(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top.
\section{Calculation of $\boldsymbol\Xi_{0}$ and $\boldsymbol\Xi_{1}$}\label{ap:Xiderivation}
The expressions for $\boldsymbol\Xi_{0}$ and $\boldsymbol\Xi_{1}$ in terms of smoother outputs can be found in \cite{Shumway2000}, here we present the derivation for completeness.  The expression for  $\boldsymbol\Xi_{0}$ as defined in Eq.~\eqref{eq:MRA-Xi0} of the main text is given by
   \begin{equation}\label{eq:appXi0}
	 \boldsymbol\Xi_0=\mathbf E_{\theta'}\left[\sum_{t=0}^{T-1}\mathbf x_t\mathbf x_{t+1}^\top\right].
	We start by computing the cross-covariance matrix as

>>>>

As mentioned earlier, the E-step for state-space models is equivalent to the smoothing problem. We use the RTS smoother, given in Table~\ref{alg:MRA-RTS} for completeness, in order to infer the states of the system. The RTS smoother uses the Kalman filter \citep{Kalman1960} in forward pass (Steps~1 and 2 in Table~\ref{alg:MRA-RTS}), along with a backward iteration to calculate its outputs (Steps~3 and 4 in Table~\ref{alg:MRA-RTS}): smoothed states $\hat{\mathbf x}^b_t$ and covariances, $\mathbf P^b_t=\mathrm{cov}(\mathbf{x}_t)$. Following \cite{Gibsona2005}, an additional backwards recursion is also included (Step~5 in Table~\ref{alg:MRA-RTS})	to calculate cross-covariance matrix, $\mathbf M_t=\mathrm{cov}(\mathbf{x}_{t},\mathbf{x}_{t+1})$, which is also required in forming the lower bound to be maximized in the M-step. Note that the following quantities from the forward iteration should be stored when the RTS smoother is implemented: predicted and corrected estimates for both state estimates ($\hat{\mathbf{x}}_t^{f-}$, $\hat{\mathbf{x}}_t^{f}$) and covariance matrices ($\mathbf P_t^{f-}$,$\mathbf P_t^f$). Also, the final value of the Kalman gain, $\mathcal K_T$, should be stored in order to initialize the backward cross-covariance computation. 

@@

For the linear state-space model given by Eqs.~\eqref{eq:StateEquation} and \eqref{eq:ReducedObservationEquation}, the bound defined by Eq.~\eqref{eq:Bishopbound} is a quadratic of the form (see \ref{ap:QDerivation} for derivation) 
\begin{equation}\label{eq:MRA-QCompact}
\mathcal Q\left(\boldsymbol \theta,\boldsymbol\theta'\right)=\beta+2T_s\varsigma\left(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1\right)\boldsymbol\theta-T_s\varsigma\boldsymbol\theta^\top\boldsymbol\Upsilon\boldsymbol\theta,
\end{equation}
where $\beta$ is constant with respect to $\boldsymbol\theta$. Equation~\eqref{eq:MRA-QCompact} is maximum at
\begin{align}\label{eq:MRA-thetahat}
\boldsymbol \theta= \boldsymbol\Upsilon^{-\top}(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top,
\end{align}
where
\begin{align}
\boldsymbol\upsilon_0&=\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_0]_{i,j}[\boldsymbol\Gamma_1]_{j,i}\label{eq:epsilon0} \\
\boldsymbol\upsilon_1&=\xi\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\boldsymbol\Gamma_1]^{j,i} \label{eq:epsilon1}\\
\boldsymbol\Upsilon&=T_s\varsigma\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\boldsymbol\Gamma_2]^{j,i},\label{eq:Epsilon}
\end{align} 
where $[\cdot]_{p,q}$ denotes the $\left(p,q\right)$-element of the matrix, and $ [\cdot]^{p,q}$ denotes the $\left(p,q\right)$-block of the block matrix and where 
\begin{align}
\left[ \Gamma_1\right]^{j,i} &=\sum_{k=1}^{n_x}\left[ \boldsymbol\Sigma_w^{-1}\boldsymbol\Lambda_x^{-1}\right]_{jk} \left[ \mathbf U\right]^{k,i},\\
\left[ \Gamma_2\right] ^{j,i}&=\sum_{k,m=1}^{n_x}[\mathbf U^{\top}]^{jk} \left[\boldsymbol\Lambda_x^{-1}\boldsymbol\Sigma_w^{-1}\boldsymbol\Lambda_x^{-1} \right]_{km}[\mathbf U]^{mi}.
\end{align}
Each $ 1 \times n_{\theta}$ block of the $n_x \times n_x n_{\theta}$ block matrix $\mathbf U$ is 
\begin{align}\label{eq:Uij}
\left[ \mathbf U\right] ^{k,i}&=\int_{\boldsymbol \Omega}\left[\boldsymbol\mu(r) \right]_k \left[\int_{\boldsymbol\Omega} \boldsymbol\mu\left(r'\right)\boldsymbol \lambda^\top \left(r-r'\right) \mathrm{d}r'\right]_{i:} \mathrm{d}r,
\end{align}
where $[\cdot]_{i:} $ denotes the $i^{th}$ row. The matrices $\boldsymbol\Xi_0$ and $\boldsymbol\Xi_1$ in Eqs.~\eqref{eq:epsilon0}-\eqref{eq:Epsilon} are calculated using the RTS smoother outputs: $\hat{\mathbf x}_t$, covariance, $\mathbf P_t^b$, and cross-covariance matrix, $\mathbf M_t$ \citep{Gibsona2005}, 
\begin{align}
\boldsymbol\Xi_0&=\sum_{t=0}^{T-1}\left(\mathbf M_{t+1}+\mathbf{\hat x}_t\mathbf{\hat x}_{t+1}^\top\right) \label{eq:Xi0} \\
 \boldsymbol\Xi_1&=\sum_{t=0}^{T-1}\left(\mathbf P_t+\mathbf{\hat x}_t\mathbf{\hat x}_t^\top\right).  \label{eq:Xi1}
\end{align} 
The second derivative of the $\mathcal Q$-function is given by
\begin{equation}
\frac{\partial^2\mathcal Q}{\partial\boldsymbol\theta^2}=-2T_s\varsigma\boldsymbol\Upsilon.
\end{equation}
The matrix $\boldsymbol\Upsilon$ is positive definite if $n_x^2\times n_{\theta}$ matrix $\mathrm{vec}(\mathbf U)$ is of rank $n_{\theta}$ (\citep{Dewar2009}, Lemma 3), where $\mathrm{vec}(\cdot)$ denotes the vectorization operator. Under this condition $\boldsymbol\Upsilon$ is  invertible and the second derivative is negative definite representing a maximum of the $\mathcal Q$-function.

\subsection{Stopping rule}
The stopping criterion is usually associated with either the change in the parameter estimates or the log-likelihood variation \citep{McLachlan1997}. The stopping rule adopted herein is
\begin{equation}
 \left(\parallel \mathbf{A} \parallel_{F}^{(i)}-\parallel \mathbf{A} \parallel_{F}^{(i-1)}\right)<\epsilon,
 \end{equation}
 where $\epsilon$ is a threshold value and $\parallel \mathbf{A} \parallel_{F}^{(i)}$ and $ \parallel \mathbf{A} \parallel_{F}^{(i-1)}$ are the Frobenius norms  of the successive estimates of $\mathbf{A} $ matrices which is defined by \citet{Meyer2000}
 \begin{equation}
  \parallel \mathbf{A} \parallel_{F}=\sqrt{\sum_{i,j=1}^{n_x}\mid a_{i,j} \mid^2}=\sqrt{\mathrm{tr} (\mathbf A^{\top}\mathbf A)}.
 \end{equation}
\subsection{Computational complexity} 
All $n_x^2$ blocks of $\mathbf U$, $\boldsymbol\Gamma_1$ and $\boldsymbol\Gamma_2$  can be computed as a one-off before the commencement of the EM iterations, which increases the speed of the M-step significantly compared to the implementation of \citet{Dewar2009}. The complexity of each step of the algorithm is summarized in Table~\ref{table:MRA-ComputationalComplexity} where for ease of comparison the computational complexity of the method of \citet{Dewar2009} is also identified. It can be seen from the table that the state dimension is an important factor at each iteration of the estimation algorithm, emphasizing the advantage of the proposed algorithm compared to the method proposed in \citet{Dewar2009}. This also suggests the use of an efficient algorithm to choose the minimum number of basis functions (states of the system) while achieving the desired details of the underlying field.  
\begin {table}
\begin{center}
\scalebox{1}{\begin{tabular}{lllc}
\hline \hline
Variable &Equation&Order&Order obtained from \citet{Dewar2009}\\ \hline\\
$\Xi_0, \Xi_1$&Eqs.~\eqref{eq:Xi0} and \eqref{eq:Xi1}&$O(Tn_x^2)$& \----\\
$\boldsymbol\upsilon_0, \boldsymbol\upsilon_1$&Eqs.~\eqref{eq:epsilon0} and \eqref{eq:epsilon1} &$O(n_x^2n_{\theta})$ &$O(n_x^2n_{\theta})+O(n_x^3)$ \\
$\boldsymbol\Upsilon$&Eq.~\eqref{eq:Epsilon}&$O(n_x^2n_{\theta}^2)$&$O(n_x^4n_{\theta}^2)$\\
$\mathbf A(\boldsymbol\theta)$&Eq.~\eqref{eq:A_theta}&$O(n_x^3)$&\----\\
$\boldsymbol\Lambda_{\theta}$&Eq.~\eqref{eq:LambdaThetaElements}&$O(n_x^2n_{\theta})$&\----\\
$\hat{\boldsymbol\theta} $&Eq.~\eqref{eq:MRA-thetahat}&$O(n_{\theta}^3)$&\----\\
\hline \hline
\end{tabular}}
 \caption {{\bf The Computational complexity of the estimation algorithm}. A comparison between the algorithm proposed herein and that of \citet{Dewar2009} is also provided. "\----'' indicate equal complexities for both algorithms.} 
\label{table:MRA-ComputationalComplexity}
\end{center}
\end {table}

@@

Each element of $\boldsymbol\Lambda_{\theta}$ can be calculated using
\begin{equation}\label{eq:LambdaThetaElements}

@@

where each $ 1 \times n_{\theta}$ block of the $n_x \times n_x n_{\theta}$ block matrix $\mathbf U$ is 
\begin{align}\label{eq:app-Uij}

@@

By substituting for $\left[ \boldsymbol\Lambda_{\theta}\right] _{k,i}$ from Eq.~\eqref{eq:app-LambdaThetaElements} the trace given in Eq.~\eqref{eq:MRA-trace1} can be written as

@@

$\boldsymbol\upsilon_0$, $\boldsymbol\upsilon_1$ and $\boldsymbol\Upsilon$ are defined in Eqs.~\eqref{eq:epsilon0}-\eqref{eq:Epsilon}.

By differentiating the $\mathcal Q$-function with respect to $\boldsymbol\theta$ we have
\begin{align}\label{eq:MRA-QDerivative}
\frac{\partial \mathcal Q}{\partial \boldsymbol\theta}&=2T_s\varsigma(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top-T_s\varsigma(\boldsymbol\Upsilon^\top+\boldsymbol\Upsilon)\boldsymbol\theta \nonumber \\
&=2T_s\varsigma(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top-2T_s\varsigma\boldsymbol\Upsilon^\top\boldsymbol\theta.
\end{align}
Equating Eq.~\eqref{eq:MRA-QDerivative} to $\mathbf 0$ vector yields
\begin{align}\label{eq:app-thetahat}
\boldsymbol \theta= \boldsymbol\Upsilon^{-\top}(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1)^\top.
\end{align}   
\section{Calculation of $\boldsymbol\Xi_{0}$ and $\boldsymbol\Xi_{1}$}\label{ap:Xiderivation}
The expressions for $\boldsymbol\Xi_{0}$ and $\boldsymbol\Xi_{1}$ in terms of smoother outputs can be found in \cite{Shumway2000}, here we present the derivation for completeness.  The expression for  $\boldsymbol\Xi_{0}$ as defined in Eq.~\eqref{eq:MRA-Xi0} of the main text is given by
   \begin{equation}\label{eq:appXi0}
	 \boldsymbol\Xi_0=\mathbf E_{\theta'}\left[\sum_{t=0}^{T-1}\mathbf x_t\mathbf x_{t+1}^\top\right].
	\end{equation}                                                                                                
	We start by computing the cross-covariance matrix as
